### 2021-06-03 memo.tex
't Hooft-Polyakov monopole argument v2

\section{'t Hooft-Polyakov monopole argument}
For an $SU(2)$ gauge theory Higgssed by an adjoint ($\bm{3}$, spin-$1$, isovector) scalar,
the gauge group is broken down to $U(1)$, and correspondingly it accommodates topological solitons (monopoles):
\begin{equation*}
	\pi_2\left(
		\dfrac{SU(2)}{U(1)}
	\right) = \bZ.
\end{equation*}
In the presence of (additional) fermions, this monopole might acquire non-trivial charge under
spacetime-Lorentz or flavor symmetries, depending on the representation of the fermions under $SU(2)$ gauge symmetry:
\begin{equation*}
	\begin{tabular}{c||cccc}
		fermion gauge rep. & number of zero-modes & spin of zero-modes& spin of monopole & flavor rep.\\
		\hline
		$\bm{2}$ & $1$ & $0$ & $0$ & spinor\\
		$\bm{3}$ & $2$ & $\tfrac{1}{2}$ & \\
		$\bm{4}$ & $4$ & $0\oplus 1$ & $\tfrac{1}{2}$\\
	\end{tabular}
\end{equation*}
The numbers of zero-modes can be computed from the Callias index theorem\cite{Callias:1977kg}.

According to \cite{Wang:2018qoy},
there is a $w_2(TM_4)\beta w_2(SU(2)_{\text{gauge}})$ anomaly for a fermion in $\bm{4}$ charged under
\begin{equation*}
	\dfrac{Spin(4)_{\text{spacetime}} \times SU(2)_{\text{gauge}}}{\bZ_2},
\end{equation*}
which incarnates in the IR as an ill-definition of the effective interaction $w_2(TM) c_1(U(1)_{\text{gauge}})$,
emerging after integrating out the fermion which obtained mass through Yukawa coupling.
This effective interaction term should arise in order to make the monopole a fermion,
but is not well-defined without a trivialization of $w_2(TM_4)$ or equivalently a spin structure.

This situation looks quite similar to our problem where the fermions are charged under
\begin{equation*}
	\dfrac{SO(4)_{\text{gauge}} \times SU(2)_{\text{flavor}}}{\bZ_2}.
\end{equation*}
Therefore, by breaking $SU(2)_{\text{flavor}}$ to $U(1)_{\text{flavor}}$
and examining the charge of the corresponding monopole under the remaining gauge group,
one might be able to find out that there is a similar problem,
which implies the existence of $w_2\beta w_2$ anomaly in the first place as desired.

Also, one should be able to generalize this whole argument to the case of fermions charged under
\begin{equation*}
	\dfrac{SO(2n_c)_{\text{gauge}} \times SU(2n_f)_{\text{flavor}}}{\bZ_2}
\end{equation*}
by breaking $SU(2n_f)_{\text{flavor}}$ to $SO(2n_f)_{\text{flavor}}$,
since there are monopoles characterized by
\begin{equation*}
	\pi_2\left(
		\dfrac{SU(2n_f)}{SO(2n_f)}
	\right)
	=
	\left\{
		\begin{array}{cl}
			\bZ_2 & (n_f \geq 2)\\
			\bZ & (n_f = 1)
		\end{array}
	\right..
\end{equation*}

>>>>

\section{'t Hooft-Polyakov monopole argument}
For an $SU(2)$ gauge theory Higgssed by an adjoint ($\bm{3}$, spin-$1$, isovector) scalar,
the gauge group is broken down to $U(1)$, and correspondingly it accommodates topological solitons (monopoles):
\begin{equation*}
	\pi_2\left(
		\dfrac{SU(2)}{U(1)}
	\right) = \bZ.
\end{equation*}
In the presence of (additional) fermions, this monopole might acquire non-trivial charge under
spacetime-Lorentz or flavor symmetries, depending on the representation of the fermions under $SU(2)$ gauge symmetry:
\begin{equation*}
	\begin{tabular}{c||cccc}
		fermion gauge rep. & number of zero-modes & spin of zero-modes& spin of monopole\\
		\hline
		$\bm{2}$ & $1$ & $0$ & $0$\\
		$\bm{3}$ & $2$ & $\tfrac{1}{2}$\\
		$\bm{4}$ & $4$ & $0\oplus 1$ & $\tfrac{1}{2}$\\
	\end{tabular}
\end{equation*}
The numbers of zero-modes can be computed from the Callias index theorem\cite{Callias:1977kg}.

According to \cite{Wang:2018qoy},
there is $w_2(TM_4)\beta w_2(SU(2)_{\text{gauge}})$ anomaly for a fermion in $\bm{4}$ charged under
\begin{equation*}
	\dfrac{Spin(4)_{\text{spacetime}} \times SU(2)_{\text{gauge}}}{\bZ_2},
\end{equation*}
which incarnates in the IR as an ill-definition of the effective interaction $w_2(TM) c_1(U(1)_{\text{gauge}})$,
emerging after integrating out the fermion which obtained mass through Yukawa coupling.
This effective interaction term should arise in order to make the monopole a fermion
(\textit{i.e.}\,spinor representation of $Spin(4)_{\text{spacetime}}$),
but is not well-defined without a trivialization of $w_2(TM_4)$ or equivalently a spin structure.

This situation looks quite similar to our problem where the fermions are charged under
\begin{equation*}
	\dfrac{SO(4)_{\text{gauge}} \times SU(2)_{\text{flavor}}}{\bZ_2}.
\end{equation*}
Since fermions are in the fundamental representation of the flavor symmetry,
breaking $SU(2)_{\text{flavor}}$ to $U(1)_{\text{flavor}}$ gives rise to
a monopole in the spinor representation this time of the $SO(4)_{\text{gauge}}$\cite{Harvey:1996ur}.
Therefore by the same logic, one can deduce that there should be
$w_2(SO(4)_{\text{gauge}})\beta w_2(SU(2)_{\text{flavor}})$ anomaly in the first place, as desired.

Also, one should be able to generalize this whole argument to the case of fermions charged under
\begin{equation*}
	\dfrac{SO(2n_c)_{\text{gauge}} \times SU(2n_f)_{\text{flavor}}}{\bZ_2}
\end{equation*}
by breaking $SU(2n_f)_{\text{flavor}}$ to $SO(2n_f)_{\text{flavor}}$,
where we have monopoles characterized by
\begin{equation*}
	\pi_2\left(
		\dfrac{SU(2n_f)}{SO(2n_f)}
	\right)
	=
	\left\{
		\begin{array}{cl}
			\bZ_2 & (n_f \geq 2)\\
			\bZ & (n_f = 1)
		\end{array}
	\right..
\end{equation*}
